---
title: "Assignment 2"
output: html_document
---

By ***Changyu Gao***, PB15121717

## 4.28

Import data
```{r}
options(digits = 3)
X1 <- read.table('T1-5.DAT')
head(X1)
solar <- X1$V2 # solar radiation column
```

Construct a Q-Q plot
```{r}
qqp <- qqnorm(solar)
qqline(solar) # add a reference line
rq <- with(qqp, cor(x, y)) # compute correlation coefficient
```

We can look up the critical value for test with $n = 40$ and $\alpha = .05$ It is $.9726 > rq=0.969$.
Therefore, we reject the hypothesis of normality at level of significance $\alpha = 0.05$.

## 4.31

Import and split data
```{r}
X2 <- read.table('T1-6.DAT')
head(X2)
splitX2 <- split(X2[,-6], X2$V6)
nonmult <- splitX2[[1]]
mult <- splitX2[[2]]
nrow(nonmult)
nrow(mult)
```

Construct qqplot and compute rqs
```{r}
plotqq <- function(col) {
  qqp <- qqnorm(col, main = "");
  qqline(col);
  with(qqp, cor(x, y))
}
layout(matrix(1:10, nrow = 2, byrow = TRUE), respect = TRUE)
par(mar=rep.int(1,4))
nonmult.rqs <-  sapply(nonmult, plotqq)
mult.rqs <-  sapply(mult, plotqq)
nonmult.rqs
mult.rqs
```
As we can see for non-multiple-sclerosis group, only V4 has a qr > 0.97 Thus we remain skeptical about the normality assumption.
For multiple-sclerosis group, V1, V3 and V5 all have qrs > 0.97 while qrs of V3 and V5 are less than 0.85.

We can use shaprio test as well to validate the normality assumption.
```{r}
results_mult <- sapply(mult, shapiro.test)
mult.shaprio_pvalues <- simplify2array(results_mult["p.value", ])
results_nonmult <- sapply(nonmult, shapiro.test)
nonmult.shaprio_pvalues <- simplify2array(results_nonmult["p.value", ])
```

According to Royston, the normality should be adequate for p-value < 0.1
```{r}
mult.shaprio_pvalues
mult.shaprio_pvalues < 0.1
nonmult.shaprio_pvalues
nonmult.shaprio_pvalues < 0.1
```

Therefore, from Shaprio test we conclude that for multiple-sclerosis group the marginal normality for the 4th variable is dubious while that for other variables is convincing, and that for non-multiple-sclerosis group the marginal normality is convincing for all variables.
