---
title: "Assignment9"
author: "Gcy"
output: html_document
---

By ***Changyu Gao***, PB15121717

## 9.20

Import data
```{r}
options(digits = 4)
X1 <- read.table("T1-5.DAT")
X1 <- X1[,c(1,2,5,6)]
n1 <- nrow(X1)
head(X1)

(meanX1 <- colMeans(X1))
covX1 <- cov(X1)
covX1

pca1 <- prcomp(X1, scale. = T)
L <- t(t(pca1$rotation) * pca1$sdev)
(L1 <- L[, 1])
(psi1 <- diag(covX1 - L1 %*% t(L1)))
(L2 <- L[, c(1,2)])
(psi2 <- diag(covX1 - L2 %*% t(L2)))
```
a) First two factor loadings are shown above.

Maximum likelihood estimates
```{r}
# m = 1
require(psych)
fa1 <- fa(covX1, n.obs = n1, fm = "ml")
loadings(fa1)
cat("Specific variances:")
(psi_fa1 <- diag(residuals(fa1))) # specific variance
cat("MLE residual:")
residuals(fa1) - diag(psi_fa1)
cat("PCA residual:")
covX1 - L1 %*% t(L1) - diag(psi1)

# m = 2
fa2 <- fa(covX1, nfactors = 2, n.obs = n1, fm = "ml")
(L2_mle <- loadings(fa2))
cat("Specific variances:")
(psi_fa2 <- diag(residuals(fa2))) # specific variance
cat("MLE residual:")
residuals(fa2) - diag(psi_fa2)
cat("PCA residual:")
covX1 - L2 %*% t(L2) - diag(psi2)
```
Factor solutions are shown above. We could compare the residual matrices between these two methods. We find that maximum likelihood method does a better job in this sense.

We can interpret the first principal component as contrasting wind with other pollution and second as maily contrasting NO2. This is different from that of using covariance matrix.

# 10.10
```{r}
r11 <- matrix(c(1, .615, .615, 1), nrow = 2)
r12 <- matrix(c(-.111, -.195, -.266, -.085), nrow = 2)
r21 <- t(r12)
r22 <- matrix(c(1, -.269, -.269, 1), nrow = 2)

r <- rbind(cbind(r11, r12), cbind(r21, r22))
r
```

```{r}
# simple matrix power using spectrum decomposition
matrix_power <- function(A, p) {
  eig <- eigen(A)
  V <- eig$vectors
  L <- eig$values
  V %*% diag(L ^ p) %*% t(V)
}

r11sqrt <- matrix_power(r11, -1/2)
r22sqrt <- matrix_power(r22, -1/2)
A <- r11sqrt %*% r12 %*% solve(r22, r21) %*% r11sqrt
B <- r22sqrt %*% r21 %*% solve(r11, r12) %*% r22sqrt
eigA <- eigen(A)
eigB <- eigen(B)
cat("Sample canonical correlations:")
sqrt(eigA$values)
cat("First canonical pair:")
eigA$vectors[, 1] %*% r11sqrt # U_1
eigB$vectors[, 1] %*% r22sqrt # U_2
```
Results are shown above.
As we can see $\hat U_1 \approx -X_1^{(1)}$ and $\hat U_2 \approx \frac 35 X_1^{(2)} + X_2^{(2)}$, $\hat U_1$ is essentially 1973 nonprimary homicides while $\hat U_2$ is some overall measurement of punishment. These two have largest correlations.
