---
title: "Assignmen8"
author: "Gcy"
output: html_document
---

By ***Changyu Gao***, PB15121717

## 9.20

Import data
```{r}
options(digits = 4)
X1 <- read.table("T1-5.DAT")
X1 <- X1[,c(1,2,5,6)]
n1 <- nrow(X1)
head(X1)

(meanX1 <- colMeans(X1))
covX1 <- cov(X1)
covX1

pca1 <- prcomp(X1)
L <- t(t(pca1$rotation) * pca1$sdev)
(L1 <- L[, 1])
(psi1 <- diag(covX1 - L1 %*% t(L1)))
(L2 <- L[, c(1,2)])
(psi2 <- diag(covX1 - L2 %*% t(L2)))
```
a) First two factor loadings are shown above.

Maximum likelihood estimates
```{r}
# m = 1
require(psych)
fa1 <- fa(covX1, covar = T, n.obs = n1, fm = "ml")
loadings(fa1)
cat("Specific variances:")
(psi_fa1 <- diag(residuals(fa1))) # specific variance
cat("MLE residual:")
residuals(fa1) - diag(psi_fa1)
cat("PCA residual:")
covX1 - L1 %*% t(L1) - diag(psi1)

# m = 2
fa2 <- fa(covX1, nfactors = 2, n.obs = n1, covar = T, fm = "ml")
(L2_mle <- loadings(fa2))
cat("Specific variances:")
(psi_fa2 <- diag(residuals(fa2))) # specific variance
cat("MLE residual:")
residuals(fa2) - diag(psi_fa2)
cat("PCA residual:")
covX1 - L2 %*% t(L2) - diag(psi2)
```
Factor solutions are shown above. We could compare the residual matrices between these two methods. We find that maximum likelihood method does a better job in this sense.

## 9.21
```{r}
# varimax solution
# pca method

pca2_r <- principal(covX1, 2, n.obs = n1, covar = T)
(L2_r <- loadings(pca2_r))
cat("Specific variances:")
(psi_pca2_r <- diag(residuals(pca2_r))) # specific variance

# Maximum likelihood method
fa2_r <- fa(covX1, nfactors = 2, covar = T, n.obs = n1, fm = "ml", rotate = "varimax")
(L2_mle_r <- loadings(fa2_r))
cat("Specific variances:")
(psi_fa2_r <- diag(residuals(fa2_r))) # specific variance
cat("MLE residual:")
residuals(fa2_r) - diag(psi_fa2_r)
cat("PCA residual:")
covX1 - L2_r %*% t(L2_r) - diag(psi_pca2_r)
```
By comparing results, we see that two solutions are consistent with each other.

## 9.22
```{r}
d <- t(X1) - meanX1
delta <- t(L2_mle) %*% diag(1/psi_fa2) %*% L2_mle
scores_mle_wls <- t(solve(delta, t(L2_mle)) %*% diag(1/psi_fa2) %*% d)
head(scores_mle_wls)
scores_mle_reg <- t(t(L2_mle) %*% solve(covX1, d))
head(scores_mle_reg)
# pca
L0 <- (t(pca1$rotation) / pca1$sdev)[c(1,2),]
pca_scores <- t(L0 %*% d)
head(pca_scores)
```
Scores are shown above. We can see that they differ quite a lot.
