---
title: "Assignment5"
output: html_document
---

By ***Changyu Gao***, PB15121717

## 6.16

Import data
```{r}
X <- read.table("T4-3.DAT")
X <- X[, -5] # only need first 4 variables
head(X)

meanX <- colMeans(X)
covX <- cov(X)
n <- nrow(X)
p <- ncol(X)
```

The outlier is shown in red.

Consider the contrast matrix and do $T^2$ test
```{r}
C <-  matrix(c(c(1,-1,0,0),
           c(1,0,-1,0),
           c(1,0,0,-1)), nrow = 3, byrow = T)
Cx <- C %*% meanX
newcov <- C %*% covX %*% t(C)
T_norm <- t(Cx)  %*% solve(newcov, Cx) # compute T^2 statistic divided by n
T_norm

fscale <- function(p, n, a) {
  p*(n-1)/(n-p)/n * qf(1-a, p, n-p)
}
alpha <- .05
threshold <- fscale(p-1, n, alpha)
threshold
```

We can see that $T^2 / n$ is 8.49, which is much greater than the threshold 0.32. Therefor we reject the null hypothesis $C\mu = 0$
at the significant level $\alpha=.05$.

We can also interpret the outlier with $x_1 = 284$ or observation 31 as a misprint and test with $x_1 = 184$ for this observation. This won't make much difference since 184 is close to the sample mean $\bar x_1$ and this change contributes little to the overall $T^2$ statistic.

For a comparison of the dynamic measurements with the static ones, we can choose the contrast vector $c = (1, 1, -1, -1)'$.

Compute 95% simultaneous interval
```{r}
c0 <- c(1, 1, -1, -1)
c0x <- c0 %*% meanX
radius <- sqrt(threshold * (c0 %*% covX %*% c0))
cat(sprintf("interval: %.2f ± %.2f", c0x, radius))
```

## 6.20

Import data
```{r}
male <- read.table('T6-11.DAT', col.names = c("tail", "wing"))
female <- read.table('T5-12.DAT', col.names = c("tail", "wing"))
head(male)
head(female)

p <- 2

with(male, plot(tail, wing)) # scatter plot
outlier_idx <- which(male$tail == 284)
outlier <- male[outlier_idx, ]
with(outlier, points(tail, wing, col='red', pch=16)) # color the outlier red
```

Eliminate the outlier and test for equality of mean vectors
```{r}
male_correct <- male[-outlier_idx, ]
n1 <- nrow(male_correct)
n2 <- nrow(female)
mu1 <- colMeans(male_correct)
mu2 <- colMeans(female)
S1 <- cov(male_correct)
S2 <- cov(female)
Spool <- ((n1 - 1) * S1 + (n2 - 1) * S2) / (n1 + n2 - 2)
diff_mu <- mu1 - mu2
T2 <- diff_mu %*% solve(Spool * (1/n1 + 1/n2), diff_mu)
T2
threshold <- (n1+n2-1) * fscale(p, n1+n2-1, alpha)
threshold

comb <- solve(Spool, diff_mu) # most responsible linear combination
comb
```

We can see that $T^2$ statistic is about 24.965, which is greater that the threshold 10.276. Therefore, we reject the hypothesis $H_0: \mu_1 - \mu_2 = 0$ at the significant level $\alpha=.05$.
And we find the most responsible linear combination vector $(-0.157, 0.093)$ by computing $S_{pooled}^{-1}(\bar x_1 - \bar x_2)$.

Compute comfidence region and simultaneous confidence intervals
```{r}
eigensys <- eigen(Spool)
eigensys$vectors # axis of the ellipse
radius <- sqrt(threshold * (1/n1 + 1/n2))
axesLength <- sqrt(eigensys$values) * radius
axesLength # length of axes

library(car)
plot.new()
ellipse1 <- ellipse(c(0, 0), Spool, radius, draw = F)
plot(ellipse1, type = 'l')

# compute 95% simultaneous confidence intervals
# first component (tail)
cat(sprintf("tail: %.2f ± %.2f", diff_mu[1], radius * sqrt(Spool[1,1])))
cat(sprintf("wing: %.2f ± %.2f", diff_mu[2], radius * sqrt(Spool[2,2])))
```

From the simultaneous confidence intervals, we cannot be sure if male or female birds are generally larger.
