---
title: "Assignment6"
output: html_document
---

By ***Changyu Gao***, PB15121717

## 6.21

Create data matrices
```{r}
n <- 20
p <- 4
S1 <- matrix(c(.459, .254, -.026, -.244,
               .254, 27.465, -.589, -.267,
               -.026, -.589, .030, .102,
               -.244, -.267, .102, 6.854), nrow = 4)
S2 <- matrix(c(.944, -.089, .002, -.719,
               -.089, 16.432, -.400, 19.044,
               .002, -.4, .024, -.094,
               -.719, 19.044, -.094, 61.854), nrow = 4)
Spool <- matrix(c(.701, .083, -.012, -.481,
                  .083, 21.949, -.494, 9.388,
                  -.012, -.494, .027, .004,
                  -.481, 9.388, .004, 34.354), nrow = 4)
x1 <- c(2.287, 12.6, .347, 14.83)
x2 <- c(2.404, 7.155, .524, 12.840)
```

a) We notice that S1 and S2 differ a lot and n1, n2 are not large. Therefore, the pooling procedure may not work very well.

Test for equality of mean vectors
```{r}
d <- x1 - x2 # difference of mean vectors
T2 <- d %*% solve(2/n * Spool, d)
T2
alpha <- .05
fscale <- function(p, n, a) {
  p*(n-1)/(n-p)/n * qf(1-a, p, n-p)
}
threshold <- (2*n-1) * fscale(p, 2*n-1, alpha) # critical value
threshold
comb <- solve(Spool, d)
comb

comb0 <- comb * sqrt(diag(Spool))
comb0
```
b) Here the $T^2$ statistic is greater than the critical value. Hence, we reject $H_0: \mu_1 - \mu_2 = 0$ at the significant level $\alpha=.05$.
c) And we find the most responsible linear combination vector $\bar a=(-.24, .16, -3.7, .01)$ by computing $S_{pooled}^{-1}(\bar x_1 - \bar x_2)$.
d) We compute the standardized vector of $\bar a$ and get $\bar a=(-.20, .75, -.61, .07)$. The second component has the largest magnitude, implying the long-term interest rate might be useful to classify a bond.

Using approximations when covariance matrices are unequal
```{r}
I1 <- 1/n * solve((S1 + S2) / n , S1)
I2 <- 1/n * solve((S1 + S2) / n , S2)
tr <- function(A) {
  if (ncol(A) == nrow(A)) sum(diag(A))
} # trace function
v <- (p + p^2) * n/ (tr(I1 %*% I1) + tr(I1) ^ 2 +
                  tr(I2 %*% I2) + tr(I2) ^ 2)
threshold <- v*p / (v-p+1) * qf(1-alpha, p, v-p+1)
threshold
```
e) We compute the critical value and find it very close to that in b) (difference < .05). Since $T^2$ statistic is greater than the critical value, we reject $H_0: \mu_1 - \mu_2 = 0$ at the significant level $\alpha=.05$. Our conclusion doesn't change.

## 8.10
Import data
```{r}
X <- read.table("T8-4.DAT")
head(X)
S <- cov(X)
S
pca <- prcomp(X)
n1 <- nrow(X)
summary(pca)
eigval <- pca$sdev ^ 2 # eigenvalues
eigval
pca$rotation # show eigenvectors, i.e. principal components
```
a) Covariance matrix and principal components are shown above.
b) From the summary we see that the first three components account for 89.9% of total variance

Compute Bonferroni simultaneous 90% confidence intervals
```{r}
alpha1 <- 0.1
fac <- qnorm(1-alpha1/(2*3)) * sqrt(2/n1)
eigval[1:3] / (1+fac) # lower end
eigval[1:3] / (1-fac) # upper end
```

c) Corresponding confidence intervals are shown above.
```{r}
screeplot(pca, type = "lines")
```

d) We plot the screeplot and decide that the data can be summarized with the first three principal components which altogether contribute to ~90% of the total variance.
